---
title: "Welcome to Introduction to Bayesian Statistics (MATH 429/629)"
author: "AG Schissler"
date: "01/23/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(scales)
library(tidyverse)
```

## Start-of-class work (5 min)

Please write a few sentences answering "What are my goal(s) for this course?"

## I. Opening activity: Where are all the cancers? (20 min)

### Posterior simulations vizualization

### Frequentist vs. Bayesaian

### Gelman opening activity?

## II. Our community (20 min)

### Students: Undergraduate and graduate students (mostly) in math, engineering, and computer science
 
- Not much programming knowledge assumed
- Some statistics knowledge assumed

### Grader: Diana Powers

Ask Diana to prepare something...

### Instructor: A. Grant Schissler, PhD

- background
- education
- research interests
- other interests

### Reciprocal interview

Questions for you:
1) What are your goals for the course?
2) What is your best outcome imagined for the course?
3) What is your worst outcome imagined for the course?
4) What is the most likely outcome imagined for the course?
5) What can I do to help you?

Questions for me/Diana?

### Take away

Assume good will among everyone here. **Ask Questions** and **Work Together** --- these are the most important reasons for class time.

## III. Course expectations and learning outcomes (20 min)

### Course structure

- Focus is on core Bayesian statistical ideas and techniques, both computational and analytic.
- Class will be cumulative, so keep up with the material and assignments!
- There will be both online instruction.
- I'll use a variety of instructional in-class activites: lecture, quick activities, more involved "lab" assignments, and student-led presentations.
- You should budget 9 hours a week for this class (2.5 in class and 6.5 outside). The bulk of your learning with happen outside of class!

2.5 hours

```{r}
inclass_dat <- read_excel("~/OneDrive - University of Nevada, Reno/Teaching/Bayes_course/git_public/course_materials/Spring2019_inclass_schedule.xlsx")
str(inclass_dat)

sum_dat <- inclass_dat %>%
  group_by(Activity) %>%
  summarise(Minutes = sum(Minutes), n = n())

sum_dat$Percent <- sum_dat$Minutes/sum(sum_dat$Minutes) * 100

## sum(sum_dat$Percent)

bp <- ggplot(sum_dat, aes(x="", y=Percent, fill=Activity)) +
    geom_bar(width = 1, stat = "identity")
## bp

pie <- bp + coord_polar("y", start=0)

pie + geom_text(aes(y = Percent/3 + c(0, cumsum(Percent)[-length(Percent)]), label = percent(Percent/100)), size=5) +
    scale_fill_brewer(palette="Blues") +
    theme_minimal()

```

```{r}
outclass_dat <- read_excel("~/OneDrive - University of Nevada, Reno/Teaching/Bayes_course/git_public/course_materials/Spring2019_outside_class_schedule.xlsx")
## str(outclass_dat)

## weekly expectation
## sum_dat <- outclass_dat %>% group_by(Week) %>% summarise(Minutes = sum(Minutes), n = n())
## mean(sum_dat$Minutes)/60

sum_dat <- outclass_dat %>%
  group_by(Activity) %>%
  summarise(Minutes = sum(Minutes), n = n())

sum_dat$Percent <- sum_dat$Minutes/sum(sum_dat$Minutes) * 100

## sum(sum_dat$Percent)

bp <- ggplot(sum_dat, aes(x="", y=Percent, fill=Activity)) +
    geom_bar(width = 1, stat = "identity")
## bp

pie <- bp + coord_polar("y", start=0)

pie + geom_text(aes(y = Percent/3 + c(0, cumsum(Percent)[-length(Percent)]), label = percent(Percent/100)), size=5) +
    scale_fill_brewer(palette="Blues") +
    theme_minimal()


```

### Syllabus

Think-pair-share

1. Please read the syllabus and list/mark 5 most important things.
2. Pair up and share your thought with your neighbor(s).
3. Share with the whole group.

<insert point vs time plot>

### More detailed learning outcomes

- course assessment matrix

### Point-based grading system

Scatterplot
<points vs meeting>

### Course resources

#### Website

[course website](http://www.grantschissler.com/teaching/SP19/STAT429/).

#### Canvas (WebCampus)

- Course announcements
- Official assignments and due dates
- Gradebook management
- Secure file submissions
- Secure file sharing
- Another way to message the instructor

#### DataCamp for online modules

#### Github

## V. Closing (10 min)

What to do here?
