---
title: "Bayesian Inference for Simple Linear Regression (Ch.14)"
author: "AG Schissler"
date: "04/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<bring to class: laptop>

## How well does the first midterm score predict the third midterm score? (5 - 10 min)

```{r}
### load scrambled midterm one and midterm three scores
reg_data <- readRDS("exam_reg_data.rds")

### explore numerically
head(reg_data)
summary(reg_data)
cor(reg_data)
var(reg_data)

## visualize relationship
ggplot2::qplot(x = MidtermOne, y = MidtermThree, data = reg_data)
```

## I. Prior specification in Bayes Theorem for the regression model (14.4 part II; 20 - 25 min)

### Joint prior for $\beta$ and $\alpha_{\overline{x}}$

### The role of standardization in prior specification

### Community of priors

- Sketical view
- Optimistic view
- Reference
- Vague/noninformative
- Weakly informative

### Apply to exam data

## II. The joint posterior for $\beta$ and $\alpha_{\overline{x}}$ (14.5; 15 - 20 min)

### Walk through p.295. 

### Provide updating rules.

### Example 14.2

### Our data

## III. Credible intervals and hypothesis testing (14.5; 10 - 15 min)

### Bayesian credible interval around the slope
### Frequentist credible interval around the slope
### Testing One-sided hypothesis about slope
### Testing Two-sided hypothesis about slope

### Our data

## IV. Predictive ditribution for the regression model (14.5; 10 - 15 min)

### Finding the predictive distribution

### predictive distribution parameters

### Our data

## Closing (5 min)

- Write down your own answer to the guiding question, "How well does Midterm One scores predict Midterm Three?"
- Share.
