---
title: "BAYESIAN INFERENCE FOR MULTIPLE LINEAR REGRESSION"
subtitle: "Chapter 19"
author: "AG Schissler"
date: "26 Apr 2021 (*updated: `r Sys.Date()`)*"
output:
  ioslides_presentation:
    widescreen: true
    smaller: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(printr)
library(Bolstad)
set.seed(04262021)
## xaringan::inf_mr('./23_ch19_multLinReg.Rmd')
```

# i. Admin & Startup

## Week 14 of 16 Focus

- Students begin work on Final Exam. 
- Read Ch.19 Multiple Linear Reg.
- Practice on HW 12 Ch.19 1-5, due Friday Apr 30.
- Receive and act on HW 11 Ch.18 1-3 feedback.

## Today's plan

- Discuss Ch.17,18 HW general feedback
- Discuss final exam expectations.
- Discuss Ch.19
- End discussion via HW 12 Ch.19 1-5 hints.

*Please contribute questions and comments*

## Discuss Ch.17 HW

17.3. It was a 2-sample t-test!

## Final Exam: Application.

- I've assessed your skills on each topic via HW and 3 midterms.
- Let's discuss the rubric and guidelines (located in Canvas).
- Now I'd like you to apply your knowledge/skills to produce a scientific poster.
- You may use powerpoint but it is possible using using `rmarkdown`. See `postr` or `posterdown`. See [https://gist.github.com/Pakillo/4854e5d760351206084f6be8abe476b2](https://gist.github.com/Pakillo/4854e5d760351206084f6be8abe476b2).
- Let's look at a model poster: alex_k_bayes_freq_presentation.pdf. I would say it has a little too much prose for a poster. Bullet points are good. Nice viz and model.
- Keep it simple, your project doesn't need to be complicated like the 646-level poster above. 
- Select or generate your data right away! Talk to me for guidance.

# Bayesian Inference for Multiple Normal Linear Regression Model (Ch. 19)

## Discuss Ch.19 content

- etext
- Example 19.1

## Using `bayes.lm`

```{r bayes.lm}
example(bayes.lm)
```

## Final HW: Ch.19 1-5 

- Execute the code and think/understand the problem for 19.1-19.3. Present, nterpret and discuss the results.
- 19.4 description is lacking. The goal is to model `bwt` as a function of the all predictors (covariates). Use the formula `bwt ~ .`. Note that the prior expects a list. See the example in `bayes.lm` for example use. Don't forget there is a prior for the intercept $\beta_0$!. **Interpret the output a little for me*.
- 19.5 Use chi-squared test from Ch.18 (see 18.3 part iii for an example). Use `post.mean` and `post.var`.

## Closing (5 min)

- What are some issues in HW that you could use guidance on?
