%A simple LaTeX syllabus template by Luke Smith.
%I believe this is a variation of a template I got by Mike Hammond...?

%\documentclass[11pt,twocolumn]{article}
\documentclass[11pt,onecolumn]{article}
\usepackage{hyperref}
\usepackage[margin=0.75in]{geometry}
\usepackage{titlesec}
\usepackage{longtable}
\usepackage{gb4e}
%\usepackage[margin=.75in]{geometry}
\usepackage{textcomp}
\pagenumbering{gobble}
%% \usepackage[utf8]{inputenc}
%% \usepackage[english]{babel}
%% \usepackage[usenames, dvipsnames]{color}

%paragraph formatting
\setlength{\parindent}{0pt}
\setlength{\parskip}{3pt}

\usepackage{url}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=blue,
}

\titlespacing\subsection{0in}{\parskip}{\parskip}
\titlespacing\section{0in}{\parskip}{\parskip}

%Just fill these in to fill in the basic syllabus information.
\newcommand{\coursename}{Introduction to Bayesian Statistics - STAT 4XX/6XX}
\newcommand{\semester}{Spring 2019}
\newcommand{\roomnumb}{AB 635}
\newcommand{\classtimes}{Mon,Wed 1:00pm - 2:15pm}
\newcommand{\myname}{A.~Grant Schissler}
%\newcommand{\myemail}{\href{mailto:aschissler@unr.edu}{aschissler@unr.edu}}
\newcommand{\myemail}{aschissler@unr.edu}
\newcommand{\office}{DMSC 224}
\newcommand{\officehours}{Tue 2:30pm-3:30pm, Wed 1:30pm-2:30pm, or by appointment}
% \newcommand{\university}{The University of Nevada, Reno}
\renewcommand{\familydefault}{\sfdefault}

\title{\textbf{\coursename}}
% \author{{\university}---{\semester}---{\roomnumb}---{\classtimes}}
\author{{\semester}---{\roomnumb}---{\classtimes}}
\date{}


\begin{document}
\maketitle

\vspace{-0.25in}
\noindent\makebox[\linewidth]{\rule{\textwidth}{1pt}}

\begin{center}
\begin{tabular}{llll}
\textbf{Instructor}:&\myname & \textbf{Contact}:&\href{mailto:\myemail}{\myemail}, 775-784-4661 (office)\\
\textbf{Office}:&\office & \textbf{Hours}:&\officehours\\
\end{tabular}
\end{center}

\section*{Catalog Description}
Statistical inference using Bayes’ Theorem. Topics include Bayesian/frequentist comparison, posterior analysis for continuous and discrete random variables, prior specification, Bayesian regression, multivariate inference, and posterior sampling through Markov Chain Monte Carlo.

\section*{Broad student learning outcomes}

\begin{description}
\item[UG1] Students will be able to demonstrate understanding of the concepts that underlie Bayesian inference and compare the results to frequentist alternatives.
\item[UG2] Students will be able to conduct Bayesian inference analytically and interpret the results.
\item[UG3] Students will be able to perform a Bayesian analysis using professional statistical packages (e.g., \textsf{Minitab}, \textsf{R}, and \textsf{Stan}).
\item[GRAD1] Students will be able to synthesize course concepts to apply Bayesian modeling techniques to real-world data in the pursuit of scientific inquiry.
\end{description}

\section*{Course outcomes}

Students will be able to $\ldots$

\begin{enumerate}
  \itemsep0em 
\item explain the role of statistics in science using sampling theory.
\item describe and summarize data numerically and visually.
\item apply probability theory to analyze uncertainty in real world problems.
\item model parameters and data using discrete random variables.
\item conduct Bayesian inference for parameters of discrete random variables.
\item model parameters and data using continuous random variables.
\item conduct Bayesian inference for a binomial proportion.
\item compare Frequentist/Bayes approaches for inferring a binomial proportion.
\item conduct Bayesian inference for a Normal mean parameter.
\item compare Frequentist/Bayes approaches for inferring a Normal mean parameter.
\item estimate parameters in a Bayesian linear regression.
\item conduct Bayesian inference for a Normal variance parameter.
\item will apply robust Bayes methods for prior misspecification.
\item conduct Bayesian inference for parameters in a multivariate Normal.
\item use Markov Chain Monte Carlo (MCMC) to sample from posterior distributions.
\end{enumerate}


\end{document}