---
title: "STAT 446/646: Introduction to Bayesian Statistics"
semester: "Spring 2021"
instructor: 'Alfred "Grant" Schissler, Ph.D.'
left_col:
  - field: "Email"
    value: "aschissler@unr.edu"
  - field: "Office Phone"
    value: "(775) 784-4661"
  - field: "Office Location"
    value: "DMS 224"
  - field: "Office Hours"
    value: "TBD"
  
right_col:
  - field: "Web"
    value: "<a href='https://www.grantschissler.com'>www.grantschissler.com</a>"
  - field: "Class Location"
    value: "Online (Zoom)"
  - field: "Class Hours"
    value: "MoWe 1:00pm to 2:15pm"
output: 
  html_document:
    css: style.css
    template: template.html
    self_contained: true
    toc: true
    toc_depth: 1
---

```{r setup, include=FALSE}
library(tidyverse)
mon <- as.Date("2021-01-25")
# fri <- as.Date("2016-01-08")
advdate <- function(obj, adv) {
 tmon <- obj + 7*(adv-1)
 tfri <- obj + 4 + 7*(adv-1)
 tmon <- format(tmon, format="%m/%d")
 tfri <- format(tfri, format="%m/%d")
 zadv <- sprintf("%02d", adv)
 tmp <- paste("Week ",zadv,sep='',", ", tmon," - ",tfri)
 return(tmp)
}
```

## Course Description

This course introduces Bayesian statistics to a technical audience. Early on, we focus on the logic behind Bayesian statistics, including the mathematical formalization of updating beliefs under uncertainty. Then we describe Bayesian methods for well-known probability models and compare these methods to classical (frequentist) alternatives. Topics include Bayes' Theorem, prior specification, Bayesian inference for discrete and continuous univariate random variables, and linear regression. We'll implement methods analytically and using software such as Minitab, R, and Stan.

The latter portions of the course focus on implementing more advanced Bayesian methods through modern computational techniques to sample posterior distributions including Markov Chain Monte Carlo (MCMC) and Hamiltonian Monte Carlo (HMC). The aim will be to provide the concepts and technical skills to conduct a realistic Bayesian data analysis. 

*Instructor's role*

- Help explain and guide student understanding and skill development.
- Organize the course structure and assignments.
- Lead weekly class discussions.
- Evaluate student work and provide feedback.

*Student's role*

- Complete the readings before the week's discussion meeting.
- Submit completed assignments on time.
- Participate in discussions. 
- Prepare for and complete course exams.
- Conduct and report on a research project (646).

## Catalog Description

Statistical inference using Bayes' Theorem. Topics include posterior analysis for continuous and discrete random variables, prior specification, Bayesian regression, multivariate inference, and posterior sampling through Markov Chain Monte Carlo.

## Course Prerequisites

STAT 352 or STAT 467/667 or with instructor approval. STAT 445/645 is a suggested preparation.

## Student Learning Outcomes

Upon successful completion of the course, students will be able to 

* demonstrate understanding of the concepts that underlie Bayesian inference and compare the results to frequentist alternatives.
* conduct Bayesian inference analytically and interpret the results.
* perfom Bayesian analyses using professional statistical packages (`R` and `Stan`).

In addition to the above 400-level outcomes, graduate students will be able to 

* synthesize course concepts to apply Bayesian modeling techniques to real-world data.

## Textbooks & Course Materials

*Introduction to Bayesian Statistics*, 3rd edition, by William M. BOLSTAD and James M. CURRAN.

[Textbook website](https://www.wiley.com/en-us/Introduction+to+Bayesian+Statistics%2C+3rd+Edition-p-9781118091562)

**MAKE SURE TO GET THE 3rd EDITION**

*Supplemental texts*

Modern best practices for coding and reporting `R` analyses:  
*R for data science: import, tidy, transform, visualize, and model data*, Hadley WICKHAM and Garrett GROLEMUND.
[Online textbook](https://r4ds.had.co.nz/)

More Bayesian intuition and applied problems:  
*Statistical Rethinking*, 2nd edition, Richard MCELREATH
[Textbook website](https://xcelab.net/rm/statistical-rethinking/)

Standard Bayesian reference textbook:  
*Bayesian Data Analysis*, 3rd edition by Andrew GELMAN et al.
[Textbook website](http://www.stat.columbia.edu/~gelman/book/)

*Mastering `R` and `RStudio`*

https://education.rstudio.com/learn/beginner/

https://education.rstudio.com/learn/intermediate/

https://education.rstudio.com/learn/expert/


# Attendance

Students are expected to attend class.
If important circumstances prevent this, it is the student's responsibility to find out what was covered in class, what was assigned for reading or homework, and what special announcements (if any) were made.
"Excessive absence" in this class will be construed to be absence from more than 10 percent of the scheduled class sessions, whether excused or unexcused.

# Assignments

## Practice exercises

"Practice" exercises will be assigned approximately weekly. Students will submit an html document and with its `rmarkdown` source in Canvas. You are encouraged to discuss assignments between each other and with instructor. However, the assignment must be completed and submitted individually. We may divide the practice between the students to discuss solutions for all the problems.

## Midterms
No Midterm.

## Final exam
No final exam.

## Exam policy
N/A.

## Makeup, Late Policy
No make-up or late assignments.
Two dropped assignments and participation grades.

## Projects
An optional project may be conducted by students to receive instructor guidance and feedback. It is not required, but hopefully would be helpful and motivating for some students.

## Participation
Participation is the most important part of this course as it required independent study. I'll drop two participation grades, including missed classes (counts as 0 regardless of the reason).

I'll assign points for each class meeting based on the following rubric:
- (1/3 of total points) attend and participate in discussions
- (1/3 of total points)  watch videos ahead of time
- (1/3 of total points) read ahead of time

# Grading

We'll use a simple weighted scheme for grading.

```{r gradeWeights, echo=F, results='asis'}
gradeDat <- data.frame( Weight = c( '60%', '40%', '00%'),
                       Drop = c( 2, 2, 0),
                       row.names = c( 'Practice Rmd', 'Participation', 'Project' ) )
knitr::kable( gradeDat, caption = 'Grading scheme' )
```

# Additional statements

## Diversity Statement

The University of Nevada, Reno is committed to providing a safe learning and work environment for all. If you believe you have experienced discrimination, sexual harassment, sexual assault, domestic/dating violence, or stalking, whether on or off campus, or need information related to immigration concerns, please contact the University’s Equal Opportunity & Title IX Office at (775) 784-1547. Resources and interim measures are available to assist you. For more information, please visit [http:www.unr.edu/equal-opportunity-title-ix](http:www.unr.edu/equal-opportunity-title-ix).

## Disability Statement
Any student with a disability needing academic adjustments or accommodations is requested to speak with the [http:www.unr.edu/drc](Disability Resource Center) as soon as possible to arrange for appropriate accommodations.

## Academic Conduct
No laptops, cell phones, mp3 players, or other electronics are to be used for personal reasons in class. If you are being disruptive during class you will be asked to leave. Disruptions in this context include inadequate participation. You must come to class on time and stay until the end of lecture. Tardy students will not be admitted to class. Please visit [http:www.unr.edu/student-conduct](http:www.unr.edu/student-conduct) for our official student code of conduct.

## Academic Success Services
Your student fees cover usage of the University Math Center, University Tutoring Center, and University Writing Center. These centers support your classroom learning; it is your responsibility to take advantage of their services. Keep in mind that seeking help outside of class is the sign of a responsible and successful student.

## University Recording Policy
Surreptitious or covert videotaping of class or unauthorized audio recording of class is prohibited by law and by Board of Regents policy. This class may be videotaped or audio recorded only with the written permission of the instructor. In order to accommodate students with disabilities, some students may have been given permission to record class lectures and discussions. Therefore, students should understand that their comments during class may be recorded.

## Academic Dishonesty
Cheating, plagiarism, or otherwise obtaining grades under false pretenses constitutes academic dishonesty according to the code of this university. Academic dishonesty will not be tolerated and penalties can include canceling a student’s enrollment without a grade or giving an F for the assignment or for the entire course. See the University Academic Standards policy: [UAM 6,502](https://www.unr.edu/administrative-manual/6000-6999-curricula-teaching-research/instruction-research-procedures/6502-academic-standards).

# Tentative course schedule

Students please study the materials before the assigned class session.
Class readings and assignments are subject to change, contingent on mitigating circumstances and the progress we make as a class.

<div class="week">`r advdate(mon, 1)`: Organization & Intro to Bayes </div>

Readings: None.
Practice: Please complete Ch. 4 practice exercises and submit in Canvas by the end of the week.
Exams: None.
Notes: None.

*Monday topic*: Welcome to 4/646! We'll discuss the structure of the course and expectations for students/instructor.
*Wednesday topic*:

<div class="week">`r advdate(mon, 2)`: Introduction to Statistical Rethinking, Ch. 1-3</div>

The introductory content that should be familiar to the students, but with a new perspective.

*Readings*.
Please read the Preface, Ch. 1, Ch. 2, and Ch. 3 in our textbook in advance of the discussion.

*Video*.
Please watch Lecture 01 and Lecture 02 in advance of the discussion.

*Practice*.
Please complete Ch. 2 and Ch. 3 practice exercises and submit in Canvas by the end of the week.

<div class="week">`r advdate(mon, 3)`: Geocentric models Ch. 4; Wiggly Orbits Ch. 4.</div>

*Readings*.
Please read the Ch. 4 in our textbook in advance of the discussion.

*Video*.
Please watch Lecture 03 and Lecture 04 in advance of the discussion.

*Practice*.
Please complete Ch. 4 practice exercises and submit in Canvas by the end of the week.

<div class="week">`r advdate(mon, 4)`: The Many Variables & Spurious Waffles Ch. 5; The Haunted DAG & Causal Terror Ch. 6</div>

*Readings*.
Please read Ch. 5 and Ch. 6 in our textbook in advance of the discussion.

*Video*.
Please watch Lecture 05 and Lecture 06 in advance of the discussion.

*Practice*.
Please complete Ch. 5,6 practice exercises and submit in Canvas by the end of the week.

<div class="week">`r advdate(mon, 5)`: Back-door Paths Ch. 7; Ulysses' Compass Ch. 7</div>

*Readings*.
Please read Ch. 7 in our textbook in advance of the discussion.

*Video*.
Please watch Lecture 07 and Lecture 08 in advance of the discussion.

*Practice*.
Please complete Ch. 7 practice exercises and submit in Canvas by the end of the week.

<div class="week">`r advdate(mon, 6)`: Conditional Manatees Ch. 8; Markov Chain Monte Carlo Ch. 9</div>

*Readings*.
Please read Ch. 8,9 in our textbook in advance of the discussion.

*Video*.
Please watch Lecture 09 and Lecture 10 in advance of the discussion.

*Practice*.
Please complete Ch. 8,9 practice exercises and submit in Canvas by the end of the week.

<div class="week">`r advdate(mon, 7)`: Big Entropy and the GLM Ch. 10; God Spiked the Integers Ch. 11</div>

*Readings*.
Please read Ch. 10,11 in our textbook in advance of the discussion.

*Video*.
Please watch Lecture 11 and Lecture 12 in advance of the discussion.

*Practice*.
Please complete Ch. 11 practice exercises and submit in Canvas by the end of the week.

<div class="week">`r advdate(mon, 8)`: Monsters & Mixtures Ch. 12; Order Categories Ch. 12</div>

*Readings*.
Please read Ch. 12 in our textbook in advance of the discussion.

*Video*.
Please watch Lecture 13 and Lecture 14 in advance of the discussion.

*Practice*.
Please complete Ch. 12 practice exercises and submit in Canvas by the end of the week.

<div class="week">`r advdate(mon, 9)`: Models with Memory Ch. 13;  Multilevel Models 2 Ch. 13</div>

*Readings*.
Please read Ch. 13 in our textbook in advance of the discussion.

*Video*.
Please watch Lecture 15 and Lecture 16 in advance of the discussion.

*Practice*.
Please complete Ch. 13 practice exercises and submit in Canvas by the end of the week.

<div class="week">`r advdate(mon, 10)`: Adventures in Covariance Ch. 14;  Further Adventures Ch. 14</div>

*Readings*.
Please read Ch. 14 in our textbook in advance of the discussion.

*Video*.
Please watch Lecture 17 and Lecture 18 in advance of the discussion.

*Practice*.
Please complete Ch. 14 practice exercises and submit in Canvas by the end of the week.

<div class="week">`r advdate(mon, 11)`: Gaussian Processes Ch. 14; Missing Data & Other Opportunities Ch. 15</div>

*Readings*.
Please read Ch. 14, 15 in our textbook in advance of the discussion.

*Video*.
Please watch Lecture 19 and Lecture 20 in advance of the discussion.

*Practice*.
Please complete Ch. 15 practice exercises and submit in Canvas by the end of the week.

<div class="week">`r advdate(mon, 12)`: Generalized Linear Madness Ch. 16; Horoscopes</div>

*Readings*.
Please read Ch. 16,17 in our textbook in advance of the discussion.

*Video*.
None unless I record or find a lecture.

*Practice*.
Please complete Ch. 16 practice exercises and submit in Canvas by the end of the week.

<div class="week">`r advdate(mon, 13)`: Catch-up / student presentations**</div>
<div class="week">`r advdate(mon, 14)`: Catch-up / student presentations**</div>
<div class="week">`r advdate(mon, 15)`: Catch-up / student presentations**</div>
<div class="week">`r advdate(mon, 16)`: Catch-up / student presentations**</div>
<div class="week">`r advdate(mon, 17)`: NO FINAL MEETING**</div>

