\documentclass[11pt,article,landscape]{memoir}
% Copyright (C) 2013 Andrew Gainer-Dewar <andrew.gainer.dewar@gmail.com>
% This file may be distributed and/or modified under the
% conditions of the LaTeX Project Public License, either
% version 1.2 of this license or (at your option) any later
% version. The latest version of this license is in:
% http://www.latex-project.org/lppl.txt
% and version 1.2 or later is part of all distributions of[
% LaTeX version 1999/12/01 or later.

\usepackage{agd-rubric}

\rubriccourse{MATH 429/629 Introduction to Bayesian Statistics}
\rubricterm{Spring 2019}
\rubricthing{Homework 4}
\rubrictopprompt{Name}
%% \rubricbottomprompt{Total Score}

\begin{document}
\maketitle

{\large Please complete exercises Ch.7, \#1,3-5,8,9,12 in Bolstad \& Curran. I selected these problems to give deliberate practice towards mastering learning outcome 2 --- \emph{students will model parameters and data using discrete and continuous random variables.}. All exercises provide practice for calculating probabilities and properties of continuously distributed rvs, including \emph{beta, uniform, normal, gamma}. You have the option of calculating these quantities by hand (using formula/tables) or by writing R code --- just be sure to justify your work well. Also, on the midterm you'll likely have to calculate by hand so keep that in mind when practicing.}

% The argument determines the number of score buckets
\begin{rubrictable}{4}{0}
  % Add your \rubricdesc items in increasing order!

  %% Category Description
  \rubriccat{Point value label}{
    \rubricdesc{Needs Improvement}{
      
    }
    \rubricdesc{Approaching standards}{
      
    }
    \rubricdesc{Meets standards}{
      
    }
    \rubricdesc{Exceeds standards}{
      
    }
    \rubrictimes{NA}
  }
  
    %% Category 1
    \rubriccat{1.~Explanation and justification}{
    \rubricdesc{}{
      Explanation is difficult to understand and is missing several components OR was not included.
    }
    \rubricdesc{}{
      Explanation is a little difficult to understand, but includes critical components.
    }
    \rubricdesc{}{
      Explanation is clear.
    }
    \rubricdesc{}{
      Explanation is detailed and clear yet concise.
    }
    \rubrictimes{2}
  }
  
  %% Category 2
    \rubriccat{2.~Mathematical accuracy}{
    \rubricdesc{}{
      More than ~25\% of computations/statements have errors in selected problems. 
    }
    \rubricdesc{}{
      Most ~(75\%-89\%) of computations/statements are free of errors in selected problems.
    }
    \rubricdesc{}{
      Almost all ~(90\%-99\%) of computations/statements are free of errors in selected problems.
    }
    \rubricdesc{}{
      All of computations/statements are free of errors in selected problems.
    }    \rubrictimes{2}
  }

  %% Category 3
    \rubriccat{3.~Completion}{
    \rubricdesc{}{
      Several problems are not completed. 
    }
    \rubricdesc{}{
      All but two of the problems are completed.
    }
    \rubricdesc{}{
      One problem not completed.
    }
    \rubricdesc{}{
      All problems addressed.
    }
    \rubrictimes{1}
  }

    %% Category 4
    \rubriccat{4.~Neatness and organization}{
    \rubricdesc{}{
      The work appears sloppy and unorganized. It is hard to know what information goes together.
    }
    \rubricdesc{}{
      The work is presented in an organized fashion but may be hard to read at times.
    }
    \rubricdesc{}{
      The work is presented in a neat and organized fashion that is usually easy to read.
    }
    \rubricdesc{}{
      The work is presented in a neat, clear, organized fashion that is easy to read.
    }
    \rubrictimes{1}
  }

    %% Category 5
    \rubriccat{5.~Mathematical terminology and notation}{
    \rubricdesc{}{
      There is little use, or a lot of inappropriate use, of terminology and notation.
    }
    \rubricdesc{}{
      Correct terminology and notation are used, but it is sometimes not easy to understand what was done.
    }
    \rubricdesc{}{
      Correct terminology and notation are usually used, making it fairly easy to understand what was done.
    }
    \rubricdesc{}{
      Correct terminology and notation are always used, making it easy to understand what was done.
    }
    \rubrictimes{1}
  }

\end{rubrictable}

\huge Total Score (out of 20 points, 21 points possible): \hrulefill
\end{document}